Consistent Polyhedral Surrogates via Embedding

### 2021-08-02 notes/top-k.tex
top k notes in progress

We will start with $\U$ as we show it is representative in Lemma~\ref{lem:top-threshold}, and the set of vertices $\R$ for $L^{top-k}$.
Consider $\R := \left\{ \frac{|M| + k -1}{k - |H|} \mathbf{1}_H + \mathbf{1}_M : H, M \subset [n], H\cap M = \emptyset, |H| + |M| \leq k \right\}$.

\begin{claim}
	For each $p \in \mathrm{relint}\simplex$, the vertices of the expected loss $\inprod{p}{L^{top-k}(u)} = 1 - \inprod{p}{u} + \sum_{i=1}^k u_{[i]} - \inprod{p}{\min(u,\ones)}$ are denoted by $\R$.
\end{claim}
\begin{proof}
	First, observe the loss is affine in the first two terms for all $u \in \U$, and non-linearity is only introduced in the last two terms.
	Fixing a set $T$ of size $k$, and for any $u \in \U$, we will call $T_k(u)$ the top $k$ elements of $u$.
	Moreover, denote by $A^T = \{u \in \U \mid T_k(u) = T\}$ as the set of $u$ whose top $k$ elements are exactly the elements of $T$.
	By construction, we now have $\bar u$ is affine in $A^T$ for each $T$.

@@

\begin{claim}
	Fix $p \in \simplex$.
	$\hat \Gamma(p) \in \R$.
\end{claim}	
\begin{proof}
	Fix $p \in \simplex$, and take $H := |H_p|, M:= |M_p|$, and $B := |B_p|$.
	By construction, we just need to show that $H + M + B \leq k$ and we are done.

	\jessie{TODO}

@@

@@

\begin{lemma}
	Fix $p \in \simplex$.
	For small enough $\varepsilon$ and any $\hat y \in \Y$, consider $\varepsilon_{\hat y} := \varepsilon e_{\hat y}$.
	For any (small enough) $\varepsilon_{\hat y}$, the loss $\inprod{L(\hat \Gamma(p))}{p} \leq \inprod{L(\hat \Gamma(p) + \varepsilon_{\hat y})}{p}$.
\end{lemma}
\begin{proof}
	We denote $L^* := \inprod{L(\hat \Gamma(p))}{p}$.
	We proceed in cases:

@@

\jessie{I think this is clear, but I'm biased as the writer}
Since $L^{top-k}$ is polyhedral, $\inprod{L^{top-k}(\hat\Gamma(p))}{p}$ is linear in any direction $\varepsilon v$ for small enough $\varepsilon$.
Moreover, $\{\varepsilon_{\hat y}\}_{\hat y \in \Y}$ forms a basis for $\reals^n$, so for all $v \in \reals^n$, there is a unique $\lambda = \{\lambda_i\} \in \reals^n$ such that $v = \lambda_i e_i$.
Therefore, we have 
\begin{align*}
\inprod{L(\hat \Gamma(p) + \sum_i \lambda_{i} \varepsilon_{i})}{p} &= 
\sum_i \lambda_i \inprod{L(\hat \Gamma(p) + \varepsilon_{i})}{p}\\
&\geq \sum_i \lambda_i L^* \\
&= L^*
\end{align*}
This yields the following corollary.

\begin{corollary}
	$\R := \hat \Gamma(\simplex)$ is a finite representative set.
\end{corollary}

>>>>

We will start with $\U$ as we show it is representative in Lemma~\ref{lem:top-threshold}, and the set of vertices $\R$ for $L^{top-k}$.
Consider $\R := \left\{ \frac{|M| + k -1}{k - |H|} \mathbf{1}_H + \mathbf{1}_M : H, M \subset [n], H\cap M = \emptyset, |H| + |M| \leq k, |H| < k \right\}$.

\begin{claim}\label{claim:R-vertices}
	For each $p \in \mathrm{relint}\simplex$, the vertices of the expected loss $\inprod{p}{L^{top-k}(u)} = 1 - \inprod{p}{u} + \sum_{i=1}^k u_{[i]} - \inprod{p}{\min(u,\ones)}$ are denoted by $\R$.
\end{claim}
\begin{proof}
	First, observe the loss is affine in the first two terms for all $u \in \U$, and non-linearity is only introduced in the last two terms.
	Fixing a set $T$ of size $k$, and for any $u \in \U$, we will call $T_k(u)$ the top $k$ elements of $u$.
	Moreover, denote by $A^T = \{u \in \U \mid T_k(u) = T\}$ as the set of $u$ whose top $k$ elements are exactly the elements of $T$.
	By construction, we now have $\bar u$ is affine in $A^T$ for each $T$.

@@

\begin{claim}\label{claim:hat-Gamma-subset-R}
	Fix $p \in \simplex$.
	$\hat \Gamma(p) \in \R$.
\end{claim}	
\begin{proof}
	Fix $p \in \simplex$, and take $H := |H_p|, M:= |M_p|$, and $B := |B_p|$.
	By construction, we just need to show that $H + M + B \leq k$ and we are done.

	\jessie{TODO}
	We can consider $8$ cases: each corresponding to when (a) H (b) M (c) B (d) H,M (e) H,B (f) B,M (g) H,M,B are nonzero.

	(a) We immediately have at most $k-1$ outcomes $y \in \Y$ with weight strictly grater than $\frac 1 k$ since distributions with $k$ elements having weight exactly $\frac 1 k$ will not be in $H_p$, and in this case, the weight of all other elements is $0$.

	(b) Similarly, we immediately have at most $k$ elements with weight strictly greater than $\frac 1 {k+1}$.

	(c) Cannot happen, since $H_p = \emptyset \implies B_p = \{i \mid p_i \in \left(\frac{1}{k+1}, \frac{1}{k+1}\right]\} = \emptyset$.

	(d) Since $p_H > \frac H k$, we have $p_M < 1 - H_k$.
	\jessie{TODO}

	(e) Suppose $p_H = 1 - \epsilon$.
	For all $b \in B_p$, we then have $p_b > \frac{k \epsilon}{(k+1)(k-H)}$.
	By simplex constraints, we then have 
	\begin{align*}
	1 - \epsilon + \frac{Bk\epsilon}{(k+1)(k-H)}&\leq 1 \\
	\iff \frac{Bk}{(k+1)(k-H)}&\leq 1 \\
	Bk &\leq (k+1)(k-H)\\
	B &\leq \frac{k^2 + k - H - Hk}{k}\\
	B &\leq k+1-H - \frac H k
	&< k+1 - H
	\end{align*}
	Thus, $B+H < k+1$.

	(f) Again, see (c) as to why this cannot happen, since $H = 0 \implies B = 0$..

	(g)

@@

\begin{claim}
	With $k < n$, we have $\hat \Gamma(\simplex) = \R$.
	\jessie{Need to move below proof of correctness}
\end{claim}
\begin{proof}
	By Claim~\ref{claim:hat-Gamma-subset-R}, we immediately have $\hat \Gamma(\simplex) \subseteq \R$.

	Now for any sets $H,M$ such that $H, M \subseteq [n]$, $H \cap M = \emptyset$, $|H| + |M| \leq k$, and $|H| < k$ with $k <n$, consider the distribution $p_{HM} = (\frac{1}{k} + \varepsilon)\ones_{H} + (\frac{1}{k+1} + \varepsilon)\ones_M + (1 - ((\frac{|H|}{k} + |H| \varepsilon) + (\frac{|M|}{k+1} + |M| \varepsilon))) \ones_{[n] \setminus H\cup M} $.
\end{proof}

@@

\begin{lemma}\label{lem:hat-gamma-opt-one-dir}
	Fix $p \in \simplex$.
	For small enough $\varepsilon$ and any $\hat y \in \Y$, consider $\varepsilon_{\hat y} := \varepsilon e_{\hat y}$.
	For any (small enough) $\varepsilon_{\hat y}$, the loss $\inprod{L(\hat \Gamma(p))}{p} \leq \inprod{L(\hat \Gamma(p) + \varepsilon_{\hat y})}{p}$.
\end{lemma}
\begin{proof}
	We denote $L^* := \inprod{L(\hat \Gamma(p))}{p}$.
	We proceed in cases:

@@

\jessie{I think this is clear, but I'm biased as the writer}
Since $L^{top-k}$ is polyhedral, $\inprod{L^{top-k}(\hat\Gamma(p))}{p}$ is linear in any direction $\varepsilon v$ for small enough $\varepsilon$.
Moreover, $\{\varepsilon_{\hat y}\}_{\hat y \in \Y}$ forms a basis for $\reals^n$, so for all $v \in \reals^n$, there is a unique $\lambda = \{\lambda_i\} \in \reals^n$ such that $v = \lambda_i e_i$.
Therefore, we have 
\begin{align*}
\inprod{L(\hat \Gamma(p) + \sum_i \lambda_{i} \varepsilon_{i})}{p} &= 
\sum_i \lambda_i \inprod{L(\hat \Gamma(p) + \varepsilon_{i})}{p}\\
&\geq \sum_i \lambda_i L^* \\
&= L^*
\end{align*}

By Claim~\ref{claim:R-vertices}, we know exactly the regions on which $L^{top-k}$ is affine.
Since L is polyhedral and we have the regions where it is affine, in order to understand if $\hat \Gamma(\simplex)$ is representative for $L^{top-k}$, we just need to check how the loss at $\hat \Gamma(p)$ compares to the loss at vertices of these affine regions for all $p \in \simplex$.

However, since $L^{top-k}$ is convex, we actually only need to check vertices of the affine regions on which $u := \hat \Gamma(p)$ sits.  
Moreover, since $L$ is affine on each of those (convex) regions defined by some $T,S$ pair, we actually only need to check the vertices adjacent to $u$.
To justify this, consider that, for a convex region $A^{T,S}$ on which $L^{top-k}$ is affine, let $\hat \Gamma(p)$ have adjacent vertices $a_i$.
For any other vertex $v$ of $A^{T,S}$, there is a point $v_\lambda = \lambda v + (1-\lambda) \hat \Gamma(p)$ for $\lambda \in [0,1]$ such that $v_\lambda = \sum_i \lambda'_i a_i$.\jessie{Right?}

Since $L^{top-k}$ is affine on $A^{T,S}$ and $|haat \Gamma(p)$ attains better loss than its adjacent vertices, we then have 
\begin{align*}
\inprod{p}{L(\hat \Gamma(p))} &\leq \inprod{p}{L(a_i)} \,\, \forall i & \text{by Claim} \\
\implies \inprod{p}{L(\hat \Gamma(p))} &\leq \sum_i \lambda'_i \inprod{p}{L(a_i)} & \text{since $L$ affine} \\
&= \inprod{p}{\sum_i L(\lambda'_i a_i)} & \text{linearity of inner product}\\
&= \inprod{p}{L(\sum_i \lambda'_i a_i)} & \text{affineness}\\
&= \inprod{p}{L(v_\lambda)} & \text{construction of $v_\lambda$}\\
&\leq \inprod{p}{L(v)} & \text{$L$ affine and $v_\lambda$ a convex combo of $v$ and $\hat \Gamma(p)$}\\
\end{align*}

Now, we can ask what are the vertices adjacent to $u$?  They are precisely the elements of $\R$ if you moved one element of $u$ to the "next bin" (e.g., low to medium, medium to high, medium to low, high to medium).  
This is the same as considering the other equality for each $i \in S$ ($0 \leq u_i \leq 1$) or $i \in T \setminus S$ ($1 \leq u_i \leq 1 + \bar u_{-i}$.)

Since $L^{top-k}$ is affine on each of the regions $A^{T,S}$,  and in particular, between vertices, we only need to move a little in the direction of the other vertex and show an increase in loss. 
The proof of Lemma~\ref{lem:hat-gamma-opt-one-dir} enumerates all the cases we need.
This yields the following corollary.

\begin{corollary}
	$\R := \hat \Gamma(\simplex)$ is a (minimum?) finite representative set.
\end{corollary}
